---
title: "Interactive Charts"
output:
  html_document:
    toc: true
---

#### Data Visualisation with R

These data visualisation modules provide an overview of packages for interactive graphs with R. <!--more--> 

See [here](http://klein.uk/teaching/) for the course materials.

***

#### Overview

- [The rCharts package](#a)
    + [Population pyramid](#a1)
    + [Data table](#a2)
- [The googleVis package](#b)
    + [Tree map](#b1)
    + [Sankey chart](#b2)
    + [Calendar chart](#b3)
    + [Pie chart](#b4)
    + [Geo chart](#b5)
    + [Motion chart](#b6)
- [Forum](http://stackoverflow.com/tags)

For other R libraries, see [leaflet](http://rstudio.github.io/leaflet/), [htmlwidgets](http://www.htmlwidgets.org/showcase_leaflet.html), [plotly](https://plot.ly/r/) and useful tutorials on data visualisation in R at [flowingdata.com](http://flowingdata.com/).

Many of the packages below require that you have [Rtools](https://cran.r-project.org/bin/windows/Rtools/) installed in addition to [base R](https://cran.r-project.org/). I also recommend using [RStudio](https://www.rstudio.com/products/rstudio/download/) as a front end.

***

####<a name="a"></a> A. The rCharts package

The [rCharts](http://rcharts.io) package allows you to create, customize and publish interactive javascript visualizations from R using a [lattice](http://www.statmethods.net/advgraphs/trellis.html) style plotting interface.

To get started, open R and install and load the package as follows.

```{r eval=FALSE}
library(devtools)
install_github('ramnathv/rCharts')
```

Then you can start exploring the examples using pre-loaded datasets or your own data. For more examples, check out the [rCharts gallery](http://rcharts.io/gallery/).

***

#####<a name="a1"></a> A.1. Population pyramid
```{r include=FALSE}
library(knitr)
opts_chunk$set(comment = NA, results = "asis", tidy = F)
```

```{r PyramidExample, cache=TRUE}
library(rCharts)
source('https://raw.githubusercontent.com/walkerke/teaching-with-datavis/master/pyramids/rcharts_pyramids.R')

## Select population counts for Qatar in 2014
dat <- getAgeTable(country='QA', year=2014)
kable(head(dat))

n1 <- nPyramid('QA', 2014, colors = c('darkred', 'silver'))
n1$show('inline', include_assets = TRUE, cdn = TRUE)
```
<br>

[Back to top](#)

***

#####<a name="a2"></a> A.2. Data table
```{r DatatableExample, cache=TRUE}
## Load data
SchoolsGroup <- read.csv("https://raw.github.com/patilv/rChartsTutorials/master/SchoolsGroup.csv")
kable(head(SchoolsGroup))

## Load styles
tab2 <- dTable(SchoolsGroup[,-1], sPaginationType = "full_numbers")
tab2$templates$script <- "http://timelyportfolio.github.io/rCharts_dataTable/chart_customsort.html" 

## Set column names
tab2$params$table$aoColumns =
  list(
    list(sType = "string_ignore_null", sTitle = "Group 1"),
    list(sType = "string_ignore_null", sTitle = "Group 2"),
    list(sType = "string_ignore_null", sTitle = "Group 3"),
    list(sType = "string_ignore_null", sTitle = "Group 4")
  )

## Produce data table
n1 <- tab2
n1$show('inline', include_assets = TRUE, cdn = TRUE)
```

<br>
<br>

[Back to top](#)

***

####<a name="b"></a> B. The googleVis package

The [googleVis](https://cran.r-project.org/web/packages/googleVis/vignettes/googleVis_examples.html) package provides an interface to [Google Charts API](https://developers.google.com/chart/interactive/docs/gallery?hl=en), allowing users to create interactive charts based on R data frames.

```{r setOptions, message=FALSE, include=FALSE}
## Before you start install and load the package and set the output option using the `options` function.
library(googleVis)
op <- options(gvis.plot.tag='chart')
```

***

#####<a name="b1"></a> B.1. Tree map
```{r TreeExample, results='asis', tidy=FALSE, cache=TRUE}
## Load the googleVis library
library(googleVis)

## Inspect data
kable(head(Regions))

## Plot
Tree <- gvisTreeMap(data = Regions,  
                    idvar = "Region", parentvar = "Parent", 
                    sizevar = "Val", colorvar = "Fac", 
                    options=list(fontSize=16))
plot(Tree)
```
<br>

[Back to top](#)

***

#####<a name="b2"></a> B.2. Sankey chart
```{r SankeyExample, results='asis', cache=TRUE}
datSK <- data.frame(From=c(rep("A",3), rep("B", 3)),
                    To=c(rep(c("X", "Y", "Z"),2)),
                    Weight=c(5,7,6,2,9,4))
kable(head(datSK))

Sankey <- gvisSankey(data = datSK, from="From", to="To", weight="Weight",
                     options=list(
                       sankey="{link: {color: { fill: '#d799ae' } },
                            node: { color: { fill: '#a61d4c' },
                            label: { color: '#871b47' } }}"))
plot(Sankey)
```
<br>

[Back to top](#)

***

#####<a name="b3"></a> B.3. Calendar chart
```{r CalendarExample, results='asis', cache=TRUE}
kable(head(Cairo))

Cal <- gvisCalendar(data = Cairo, 
                    datevar="Date", 
                    numvar="Temp",
                    options=list(
                      title="Daily temperature in Cairo",
                      height=320,
                      calendar="{yearLabel: { fontName: 'Times-Roman',
                               fontSize: 32, color: '#1A8763', bold: true},
                               cellSize: 10,
                               cellColor: { stroke: 'red', strokeOpacity: 0.2 },
                               focusedCellColor: {stroke:'red'}}")
)
plot(Cal)
```
<br>

[Back to top](#)

***

#####<a name="b4"></a> B.4. Pie chart
```{r PieChartExample, results='asis', cache=TRUE}
kable(head(CityPopularity))

Pie <- gvisPieChart(data = CityPopularity)
plot(Pie)
```
<br>

[Back to top](#)

***

#####<a name="b5"></a> B.5. Geo chart
```{r GeoChartExample, results='asis', cache=TRUE}
kable(head(Exports))

Geo <- gvisGeoChart(data = Exports, locationvar="Country", 
                 colorvar="Profit",
                 options=list(projection="kavrayskiy-vii"))
plot(Geo)
```
<br>

[Back to top](#)

***

#####<a name="b6"></a> B.6. Motion chart
```{r MotionChartExample, results='asis', cache=TRUE}
kable(head(Fruits))

M <- gvisMotionChart(data = Fruits, idvar = 'Fruit', timevar = 'Year',
         options=list(width=400, height=350))
plot(M)
```
<br>

[Back to top](#)

***

```{r resetOptions, include=FALSE}
## At the end of your session, do not forget to reset the options.
## Set options back to original options
options(op)
```

[Back to top](#)

<a name="forum"> </a>




