---
published: true
layout: post
title: Interactive Charts
comments: true
excerpt: null
thumbnail: /viz/misc/rMaps.png
output:
  html_document:
    toc: true
---

#### Open-source data visualisations

I'm sharing some inspiring visualisations I explored with workshop participants at a datavis workshop in Ghana this week. They provide an overview of popular software packages for interactive graphs. <!--more--> 

See [here](http://klein.uk/teaching/) for the course materials.

***

#### Overview

- [The rCharts package](#a)
    + [Population pyramid](#a1)
    + [Data table](#a2)
- [The googleVis package](#b)
    + [Tree map](#b1)
    + [Sankey chart](#b2)
    + [Calendar chart](#b3)
    + [Pie chart](#b4)
    + [Geo chart](#b5)
    + [Motion chart](#b6)
- [The rMaps package](#c)
    + [Leaflet heat maps](#c1)
    + [Animated choropleths](#c2)
- [Forum](http://stackoverflow.com/tags)

For other R libraries, see [leaflet](http://rstudio.github.io/leaflet/), [htmlwidgets](http://www.htmlwidgets.org/showcase_leaflet.html), [plotly](https://plot.ly/r/) and useful tutorials on data visualisation in R at [flowingdata.com](http://flowingdata.com/).

Many of the packages below require that you have [Rtools](https://cran.r-project.org/bin/windows/Rtools/) installed in addition to [base R](https://cran.r-project.org/). I also recommend using [RStudio](https://www.rstudio.com/products/rstudio/download/) as a front end.

***

####<a name="a"></a> A. The rCharts package

The [rCharts](http://rcharts.io) package allows you to create, customize and publish interactive javascript visualizations from R using a [lattice](http://www.statmethods.net/advgraphs/trellis.html) style plotting interface.

To get started, open R and install and load the package as follows.

```{r eval=FALSE}
library(devtools)
install_github('ramnathv/rCharts')
```

Then you can start exploring the examples using pre-loaded datasets or your own data. For more examples, check out the [rCharts gallery](http://rcharts.io/gallery/).

***

#####<a name="a1"></a> A.1. Population pyramid
```{r include=FALSE}
library(knitr)
opts_chunk$set(comment = NA, results = "asis", tidy = F)
```

```{r PyramidExample, cache=TRUE}
library(rCharts)
source('http://klein.uk/R/Viz/pyramids.R')

## Select population counts for Qatar in 2014
dat <- getAgeTable2(country='QA', year=2014)
kable(head(dat))

n1 <- nPyramid(dat, colors = c('darkred', 'silver'))
n1$show('inline', include_assets = TRUE, cdn = TRUE)
```
<br>

[Back to top](#)

***

#####<a name="a2"></a> A.2. Data table
```{r DatatableExample, cache=TRUE}
## Load data
SchoolsGroup <- read.csv("https://raw.github.com/patilv/rChartsTutorials/master/SchoolsGroup.csv")
kable(head(SchoolsGroup))

## Load styles
tab2 <- dTable(SchoolsGroup[,-1], sPaginationType = "full_numbers")
tab2$templates$script <- "http://timelyportfolio.github.io/rCharts_dataTable/chart_customsort.html" 

## Set column names
tab2$params$table$aoColumns =
  list(
    list(sType = "string_ignore_null", sTitle = "Group 1"),
    list(sType = "string_ignore_null", sTitle = "Group 2"),
    list(sType = "string_ignore_null", sTitle = "Group 3"),
    list(sType = "string_ignore_null", sTitle = "Group 4")
  )

## Produce data table
n1 <- tab2
n1$show('inline', include_assets = TRUE, cdn = TRUE)
```

<br>
<br>

[Back to top](#)

***

####<a name="b"></a> B. The googleVis package

The [googleVis](https://cran.r-project.org/web/packages/googleVis/vignettes/googleVis_examples.html) package provides an interface to [Google Charts API](https://developers.google.com/chart/interactive/docs/gallery?hl=en), allowing users to create interactive charts based on R data frames.

```{r setOptions, message=FALSE, include=FALSE}
## Before you start install and load the package and set the output option using the `options` function.
library(googleVis)
op <- options(gvis.plot.tag='chart')
```

***

#####<a name="b1"></a> B.1. Tree map
```{r TreeExample, results='asis', tidy=FALSE, cache=TRUE}
## Load the googleVis library
library(googleVis)

## Inspect data
kable(head(Regions))

## Plot
Tree <- gvisTreeMap(Regions,  
                    "Region", "Parent", 
                    "Val", "Fac", 
                    options=list(fontSize=16))
plot(Tree)
```
<br>

[Back to top](#)

***

#####<a name="b2"></a> B.2. Sankey chart
```{r SankeyExample, results='asis', cache=TRUE}
datSK <- data.frame(From=c(rep("A",3), rep("B", 3)),
                    To=c(rep(c("X", "Y", "Z"),2)),
                    Weight=c(5,7,6,2,9,4))
kable(head(datSK))

Sankey <- gvisSankey(datSK, from="From", to="To", weight="Weight",
                     options=list(
                       sankey="{link: {color: { fill: '#d799ae' } },
                            node: { color: { fill: '#a61d4c' },
                            label: { color: '#871b47' } }}"))
plot(Sankey)
```
<br>

[Back to top](#)

***

#####<a name="b3"></a> B.3. Calendar chart
```{r CalendarExample, results='asis', cache=TRUE}
kable(head(Cairo))

Cal <- gvisCalendar(Cairo, 
                    datevar="Date", 
                    numvar="Temp",
                    options=list(
                      title="Daily temperature in Cairo",
                      height=320,
                      calendar="{yearLabel: { fontName: 'Times-Roman',
                               fontSize: 32, color: '#1A8763', bold: true},
                               cellSize: 10,
                               cellColor: { stroke: 'red', strokeOpacity: 0.2 },
                               focusedCellColor: {stroke:'red'}}")
)
plot(Cal)
```
<br>

[Back to top](#)

***

#####<a name="b4"></a> B.4. Pie chart
```{r PieChartExample, results='asis', cache=TRUE}
kable(head(CityPopularity))

Pie <- gvisPieChart(CityPopularity)
plot(Pie)
```
<br>

[Back to top](#)

***

#####<a name="b5"></a> B.5. Geo chart
```{r GeoChartExample, results='asis', cache=TRUE}
kable(head(Exports))

Geo=gvisGeoChart(Exports, locationvar="Country", 
                 colorvar="Profit",
                 options=list(projection="kavrayskiy-vii"))
plot(Geo)
```
<br>

[Back to top](#)

***

#####<a name="b6"></a> B.6. Motion chart
```{r MotionChartExample, results='asis', cache=TRUE}
kable(head(Fruits))

M <- gvisMotionChart(Fruits, 'Fruit', 'Year',
         options=list(width=400, height=350))
plot(M)
```
<br>

[Back to top](#)

***

```{r resetOptions, include=FALSE}
## At the end of your session, do not forget to reset the options.
## Set options back to original options
options(op)
```

[Back to top](#)

***

####<a name="c"></a> C. The rMaps package

The [rMaps](http://rmaps.github.io) package makes it easy to create, customize and share interactive maps from R, with a few lines of code. It supports several javascript based mapping libraries like Leaflet, DataMaps and Crosslet.

To get started, open R and install and load the package as follows.

```{r eval=FALSE}
library(devtools)
install_github('ramnathv/rMaps')
```

***

#####<a name="c1"></a> C.1. Leaflet heat maps
```{r MapExample, results='asis', cache=TRUE, message=FALSE}
## 1. Create map
library(rMaps)
L2 <- Leaflet$new()
L2$setView(c(29.7632836,  -95.3632715), 15)
L2$tileLayer(provider = "MapQuestOpen.OSM")
L2$show('inline', include_assets = TRUE, cdn = TRUE)

## 2. Get data
data(crime, package="ggmap")
library(plyr)
crime_dat <- ddply(crime, .(lat, lon), summarise, count = length(address))
kable(head(crime_dat))

## 3. Transform to json format
crime_dat <- rCharts::toJSONArray2(na.omit(crime_dat), json = F, names = F)

## 4. Add leaflet-heat plugin. 
L2$addAssets(jshead = c(
  "http://leaflet.github.io/Leaflet.heat/dist/leaflet-heat.js"
))

## 5. Add javascript to modify underlying chart
L2$setTemplate(afterScript = sprintf("<script> var addressPoints = %s
  var heat = L.heatLayer(addressPoints).addTo(map) </script>", 
  rjson::toJSON(crime_dat)
))
L2$show('iframesrc', cdn  = TRUE)
```
<br>

[Back to top](#)

***

#####<a name="c2"></a> C.2. Animated choropleths
```{r ChoroplethsExample, results='asis', cache=TRUE, message=FALSE}
## Get data
library(Quandl)
vcData <- Quandl("FBI_UCR/USCRIME_TYPE_VIOLENTCRIMERATE")
kable(head(vcData[,1:6]))

## Reshape data
library(reshape2)
datm <- melt(vcData, 'Year', 
  variable.name = 'State',
  value.name = 'Crime'
)
datm <- subset(na.omit(datm), 
  !(State %in% c("United States", "District of Columbia"))
)

## Discretize crime rates
datm2 <- transform(datm,
  State = state.abb[match(as.character(State), state.name)],
  fillKey = cut(Crime, quantile(Crime, seq(0, 1, 1/5)), labels = LETTERS[1:5]),
  Year = as.numeric(substr(Year, 1, 4))
)

## Associate fill colors
fills = setNames(
  c(RColorBrewer::brewer.pal(5, 'YlOrRd'), 'white'),
  c(LETTERS[1:5], 'defaultFill')
)

## Create payload for data maps
library(plyr); library(rMaps)
dat2 <- dlply(na.omit(datm2), "Year", function(x){
  y = rCharts::toJSONArray2(x, json = F)
  names(y) = lapply(y, '[[', 'State')
  return(y)
})

## Create simple choropleth
options(rcharts.cdn = TRUE)
map <- Datamaps$new()
map$set(
  dom = 'chart_1',
  scope = 'usa',
  fills = fills,
  data = dat2[[46]],
  legend = TRUE,
  labels = TRUE
)
map$show('inline', include_assets = TRUE, cdn = TRUE)
```

<br>
<br>

```{r ChoroplethsExampleContd, results='asis', cache=TRUE, message=FALSE}
## Animated choropleth
map2 = map$copy()
map2$set(
  bodyattrs = "ng-app ng-controller='rChartsCtrl'"
)
map2$addAssets(
  jshead = "http://cdnjs.cloudflare.com/ajax/libs/angular.js/1.2.1/angular.min.js"
)

map2$setTemplate(chartDiv = "
  <div class='container'>
    <input id='slider' type='range' min=2005 max=2010 ng-model='year' width=200>
    <span ng-bind='year'></span>
    <div id='{{chartId}}' class='rChart datamaps'></div>  
  </div>
  <script>
    function rChartsCtrl($scope){
      $scope.year = 2005;
      $scope.$watch('year', function(newYear){
        map{{chartId}}.updateChoropleth(chartParams.newData[newYear]);
      })
    }
  </script>"
)
map2$set(newData = dat2)
map2$show('iframesrc', cdn  = TRUE)
```
<br>

[Back to top](#)

<a name="forum"> </a>


<!--
library(knitr)
knit("~/Documents/Sites/thiloklein.github.io/viz/Rmd/2015-08-30-dataviz.Rmd", "~/Documents/Sites/thiloklein.github.io/_posts/2015-08-30-dataviz.md") 
replace ```r ``` with {% highlight r %} {% endhighlight %}
sudo jekyll serve --watch
-->


