---
title: "Choropleth maps with R"
layout: default
group: Teaching
graphics: yes
output: html_document
published: true
comments: true
---

Examples of plotting region-level data on country maps using the <a href="http://docs.ggplot2.org/current/">ggplot2</a> package and shape files from <a href="http://gadm.org">gadm.org</a>.

***

#### Overview

1. [Simulated data for Norwegian communes](#I)    
2. [Region-level literacy data from Ghanaian census](#II)


***

```{r global_options, echo=FALSE}
knitr::opts_chunk$set(fig.align = 'center', echo = TRUE, warning = FALSE, message = FALSE, error = FALSE, cache = TRUE, tidy = TRUE)
#fig.width = 5, fig.height = 4, , results = "asis", tidy = FALSE
```

Start by installing the following packages.
```{r, eval=FALSE}
## load packages
install.packages("raster")
install.packages("ggplot2")
install.packages("rgeos")
install.packages("maptools")
install.packages("scales")
```

***

#### <a name="I"></a> 1. Simulated data for Norwegian communes


```{r}
## download shapefile from gadm.org 
## level=1 -> region ; level=2 -> commune
library(raster)
states.shp <- getData('GADM', country = 'NOR', level = 2)
class(states.shp)
names(states.shp)
head(states.shp$NAME_2)
```

```{r}
## fortify shape file to dataframe format 
library(rgeos)
library(maptools)
library(ggplot2)
states.shp.f <- fortify(states.shp, region = "NAME_2")
class(states.shp.f)
head(states.shp.f)
```

```{r}
## create random data to plot on map
num.states <- length(states.shp$NAME_2)
set.seed(123)
mydata <- data.frame(id = states.shp$NAME_2, prevalence = rnorm(num.states, 55, 20))
head(mydata)
```

```{r}
## merge shape file with data 
merge.shp.coef <- merge(states.shp.f, mydata, by = "id")
head(merge.shp.coef)
```

```{r}
## basic plot
ggplot() +
  geom_polygon(data = merge.shp.coef, 
               aes(x = long, y = lat, group = group, fill = prevalence),
               color = "black", size = 0.25) + 
  coord_map()
```

```{r}
## nicer plot
library(scales)
library(ggmap)
ggplot() +
  geom_polygon(data = merge.shp.coef, 
               aes(x = long, y = lat, group = group, fill = prevalence), 
               color = "black", size = 0.25) + 
  coord_map() +
  scale_fill_distiller(name = "prevalence", palette = "YlGn", breaks = pretty_breaks(n = 5)) +
  theme_nothing(legend = TRUE) +
  labs(title = "Prevalence of X in Norway")
```

[Back to top](#)

***


#### <a name="II"></a> 2. Region-level literacy data from Ghanaian census

Now, let us try to use the above code snippets to produce regional-level plots for Ghana based on the 2010 census. Take the literacy rate as an example.

```{r}
## read individual-level literacy data from 2010 census
load(url("http://klein.uk/R/Viz/litGH.RData"))

## drop population below 11 age 
litGH <- litGH[litGH$LITERACY != "Less than 11",]

## define binary literacy variable
litGH$literacy <- ifelse(litGH$LITERACY == "None (Not literate)", 0, 1)
litGH$id <- litGH$REGION

## obtain region-level literacy rates
(litGH <- aggregate(formula = literacy ~ id, data = litGH, FUN = mean))
mydata <- litGH
```

Start over from 1. to download the shapefiles for Ghana (ISO code `'GHA'`) and produce the choropleth maps. Note: make sure to use regional levels, i.e. `level = 1` and `NAME_1`, instead of the commune-level variables in the Norwegian data above.



```{r}
## download shapefile from gadm.org 
## level=1 -> region ; level=2 -> commune
library(raster)
states.shp <- getData('GADM', country = 'GHA', level = 1)
class(states.shp)
names(states.shp)
states.shp$NAME_1
```

```{r}
## fortify shape file to dataframe format 
library(rgeos)
library(maptools)
library(ggplot2)
states.shp.f <- fortify(states.shp, region = "NAME_1")
class(states.shp.f)
head(states.shp.f)
```

```{r}
## merge shape file with data 
merge.shp.coef <- merge(states.shp.f, mydata, by = "id")
head(merge.shp.coef)
```

```{r}
## basic plot
ggplot() +
  geom_polygon(data = merge.shp.coef, 
               aes(x = long, y = lat, group = group, fill = literacy),
               color = "black", size = 0.25) + 
  coord_map()
```

```{r}
## add the region names to the basic map

## 1. obtain mean coordinates for each region
cnames <- aggregate(cbind(long, lat) ~ id, data=merge.shp.coef, FUN=function(x) mean(range(x)))

## 2. plot using geom_text option
ggplot() +
  geom_polygon(data = merge.shp.coef, 
               aes(x = long, y = lat, group = group, fill = literacy),
               color = "black", size = 0.25) + 
  coord_map() + 
  geom_text(data=cnames, aes(long, lat, label = id), size=3, color="white", fontface="bold")
```


[Back to top](#)






