---
title: "Datavis with ggplot2"
graphics: yes
output: 
  html_document
group: Teaching
published: yes
comments: yes
---

```{r global_options, echo=FALSE}
knitr::opts_chunk$set(fig.width=5, fig.height=4, fig.align='center', echo=TRUE, warning=FALSE, message=FALSE, error=FALSE, cache=TRUE)
```

```{r, echo=FALSE}
library(knitr)
library(rmarkdown)
library(scales)
library(plyr)
library(reshape2)
library(RColorBrewer)
library(ggplot2)
```


The <a href="http://docs.ggplot2.org/current/">ggplot2</a> package is a plotting system for R, based on the grammar of graphics, which tries to take the good parts of base and lattice graphics and none of the bad parts. 

<!--
For references and further examples, see:    
- [ggplot2 documention](http://ggplot2.org/book/qplot.pdf)    
- [ggplot2 website](http://blog.echen.me/2012/01/17/quick-introduction-to-ggplot2/)    
- [Edwin Chen's blog](http://blog.echen.me/2012/01/17/quick-introduction-to-ggplot2/)    
-->

***

#### Overview

1. [Introduction: Plotting with ggplot2](#I)
    + [A simple function: qplot](#a1)
    + [A robust function: ggplot](#a2)
2. [Scatterplot](#1)
3. [Faceting](#2)
     -  [Facet wrap](#2.a)
     -  [Facet grid](#2.b)
4. [Title and axis](#3)
5. [Histograms](#4)
6. [Boxplots and violin plots](#5)
7. [Output and saving](#6)

***



#### <a name="I"></a> 1. Introduction: Plotting with ggplot2

First, install and load the ggplot2 package
```{r, eval=FALSE}
install.packages("ggplot2")
library(ggplot2)
```

For this session, we will explore the `iris` data that is already pre-loaded in R.
```{r, eval=FALSE}
head(iris)
```
```{r, echo=FALSE}
kable(head(iris))
```

You can read about the data by typing
```{r, eval=FALSE}
?iris
```

<br />  


<a name="a1"></a>**1.1. A simple function: qplot **

The general formula is `qplot(x, y, ...)`. To produce a basic scatter plot, type

```{r}
qplot(x=Sepal.Length, y=Petal.Length, data=iris)   
```

Adding color per species and size depending of petals width

```{r}
qplot(Sepal.Length, Petal.Length, data=iris,
      color=Species, size=Petal.Width, 
      xlab="Sepal", ylab="Petal", main="Iris dataset")
```

We also add a title and labels for the x and y-axis, using `main`, `xlab`and `ylab`. 
       
To start at 0 for the y-axis: add `qplot(...,ylim=c(0,35))`    
This set lower and upper bounds for y axis.   
You can do the same for the x-axis

```{r}
qplot(Sepal.Length, Petal.Length, data=iris, 
      color=Species, size=Petal.Width, alpha=I(0.7), 
      xlab="Sepal Length", ylab="Petal Length", main="Iris dataset")
```
By setting the alpha of each point to 0.7, we reduce the effects of overplotting.

<br />  


<a name="a2"></a> **1.2. A robust function: ggplot**

General formula:     
`ggplot(data, aes(x,y)) + geom_*()`    
ggplot begins a plot that you finish by adding layers to, using `geom()`. ggplot provides more control than qplot().     


* *aes*: aesthetic, visual properties of the graph
     + *options aes*: color, fill, shape, size
* *geom*: graphical property 
     + geom_line; geom_bar; geom_histogram
     + geom_chart; geom_hex, geom_c(point,line) etc. You can specify for each geom the aesthetic mappings, and a default stat and position adjustment: `geom_*(aes(color=, fill=, size=....))`
* *additional elements* :
     + You can add a smoothing trend : + `geom_smooth (method="lm")` 
     + You can change the background (Themes): `+ theme_bw()`, `+theme_classic()


```{r}
ggplot(mtcars,aes(x=disp,y=mpg))+ geom_point()
```
From now on, we will only work using ggplot. 

[Back to top](#)

*** 



#### <a name="1"></a> 2. Scatterplot         
```{r}
ggplot(diamonds, aes(x=carat, y=price)) + geom_point()
```

```{r}
ggplot(diamonds, aes(x=carat, y=price, color=clarity)) + geom_point()
```
Adding color: the color of the points is determined by the clarity of the diamonds. 

```{r}
ggplot(diamonds, aes(x=carat, y=price, color=cut)) + geom_point()
```
Here we changed the color by the parameter "cut" of the dataset diamonds. 

```{r}
ggplot(diamonds, aes(x=carat, y=price, color=clarity, size=cut)) + geom_point()
```
Here we changed the size of the points according to the parameter cut. 

[Back to top](#)

*** 



#### <a name="2"></a> 3. Faceting

* Divide your plot up into multiple plots, one for each level.  
* `facet_*()` function

<a name="2.a"></a> **3.1. Facet wrap**

Write a tilde (~),  and then the attribute we would like to divide the plots by (here "clarity")
```{r}
ggplot(diamonds, aes(x=carat, y=price, color=cut)) + geom_point() + facet_wrap(~ clarity)
```

<a name="2.b"></a> **3.2. Facet grid**

To divide your graph based on **two different attributes**: `facet_grid()`       
Here we could use `"color ~ clarity"`. The tilde (~) means "explained by."

```{r}
ggplot(diamonds, aes(x=carat, y=price, color=cut)) + geom_point() + facet_grid(color ~ clarity)
```

[Back to top](#)

***



#### <a name="3"></a> 4. Title and axis

Title: add to the end of the line of code `ggtile()`  
```{r}
ggplot(diamonds, aes(carat,price)) + geom_point() + ggtitle("My scatter plot")
```

Axis label: `xlab` and `ylab`
```{r}
ggplot(diamonds, aes(x=carat, y=price)) + geom_point() + ggtitle("My scatter plot") + xlab("Weight (carats)")
```

To limit the range of the x or the y axes: `xlim` or `ylim`
```{r}
ggplot(diamonds, aes(x=carat, y=price)) + geom_point() + ggtitle("My scatter plot") + xlab("Weight (carats)") + xlim(0, 2)
```
This code limits the diamonds weight from 0 to 2 carats.     

Another possibility is to put one of the axes on a log scale. You can do this with the `scale_y_log10()` function:
```{r}
ggplot(diamonds, aes(x=carat, y=price)) + geom_point() + ggtitle("My scatter plot") + xlab("Weight (carats)") + scale_y_log10()
```

[Back to top](#)

***



#### <a name="4"></a> 5. Histograms 

Using the data set diamonds, we visualize the density of the distribution of price:
```{r}
ggplot(diamonds, aes(x=price)) + geom_histogram()
```
You notice that we do not precise the y axis in the aesthetic. Indeed, using histograms to visualize the density of distribution, y is directly equal to the density within each bin!    

You can change the width of each bin as an option, using `binwidth` inside the `geom_histogram()` function:
```{r}
ggplot(diamonds, aes(x=price)) + geom_histogram(binwidth=2000)
```

And we can add all the options that we used with a basci scatter plot. For instance, the `facet_wrap()` function:
```{r}
ggplot(diamonds, aes(x=price)) + geom_histogram(binwidth=200) + facet_wrap(~ clarity)
```

Here, all the subplots share the same y-axis: this makes it hard to interpret the frequencies...          
Some subplots have far more points than others.     

To have different y-axis for each subplot, add `scale="free_y"`
```{r}
ggplot(diamonds, aes(x=price)) + geom_histogram(binwidth=200) + facet_wrap(~ clarity, scale="free_y")
```
The same function can be used differentiating for the x-axis. 

To make a **stacked histogram** based here on the clarity attribute: Try adding the `fill` aesthetic within the aesthetic of the `ggplot()` function. 
```{r}
ggplot(diamonds, aes(x=price, fill=clarity)) + geom_histogram()
```

[Back to top](#)

***



#### <a name="5"></a> 6. Boxplots and violin plots

Boxplots are useful to compare multiple distributions. Compare the distribution of the price within each color using `geom_boxplot()`:
```{r}
ggplot(diamonds, aes(x=color, y=price)) + geom_boxplot()
```

Put the y-axis on a log scale: we get a better sense of how the distribution of price differs across multiple colors.
```{r}
ggplot(diamonds, aes(x=color, y=price)) + geom_boxplot() + scale_y_log10()
```

A box plot will always look like a square. We can instead view the distribution as a density using a "violin plot".

```{r}
ggplot(diamonds, aes(x=color, y=price)) + geom_violin() + scale_y_log10()

ggplot(diamonds, aes(x=color, y=price)) + geom_violin() + scale_y_log10() + facet_wrap(~ clarity)
```
We used the `facet_wrap()` function to display multiple subplots. 

[Back to top](#)

***



#### <a name="6"></a> 7. Output and saving     

Run your line of code and save it to a variable. For instance, call it p for plot:
```{r}
p = ggplot(diamonds, aes(x=carat, y=price)) + geom_point()
```

Save that plot to a file
```{r, eval=FALSE}
ggsave(filename="diamonds.png", p)  ## as a png
ggsave(filename="diamonds.pdf", p)  ## as a pdf
ggsave(filename="diamonds.jpeg", p) ## as a jpeg
```

One useful shortcut is that if you just displayed a plot, like in a line like this:
```{r, eval=FALSE}
ggplot(diamonds, aes(x=carat, y=price)) + geom_point()
ggsave("diamonds.png")
```
Then `ggsave()` will know to save that plot by default when you perform `ggsave` - you don't even have to tell it which plot you're saving.

[Back to top](#)

